---
title: "p8105_hw2_cy2522"
author: "Chu YU"
date: "30th-Sep-2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# problem 1
## import and clean the data
```{r}
# import the csv data
library(tidyverse)
transit_df = 
  read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names()

# clean the data
transit_tidy_df = 
  select( transit_df, line:entry, vending, entrance_type, ada) %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

# dimension
dim(transit_tidy_df)
```
This dataset contains variable: line, station_name,station_latitude
,station_longitude, route_num ,route , entry, vending, entrance_type, ada. 
My data cleaning includes three steps: cleaning variables' names, deleting the unnecessary columns and removing the missing values.
The dimension of the dataset is 20548*10. They are not tidy.There are different routes in key variables and missing values.

## distinct 
```{r}
# How many distinct stations?
sn_dis = 
  distinct(transit_tidy_df[c("station_name","line" ,"ada")]) 
nrow(sn_dis)

# How many stations with ada:
filter(sn_dis, ada == "TRUE") %>% nrow()

# prop of vending

prop.table(table(transit_tidy_df$vending))
```

## Reformat
```{r}
# gather the route columns
transit_tidy_df_rt = 
  gather(transit_tidy_df, key = route_num, value = route, route1:route11)

# how many stations serving A ?
filter(transit_tidy_df_rt, route == "A") %>% 
  distinct(station_name,line) %>%
  nrow()

# serving A as well as ADA compliance
filter(transit_tidy_df_rt, route == "A" & ada == "TRUE") %>% 
  distinct(station_name,line) %>%
  nrow()
```


# Problem 2
```{r}
# import and clean the data
trash_df = 
  readxl::read_xlsx("./data/HealthyHarborWaterWheelTotals2017-3.xlsx", 
                    range = "A2:N213", sheet = 1) %>%
  janitor::clean_names() %>%
  filter(!is.na(dumpster)) %>%
  mutate(sports_balls = round(sports_balls, digits = 0)) 

# Read and clean precipitation data for 2016 and 2017
trash_df_2017 = 
  readxl::read_xlsx("./data/HealthyHarborWaterWheelTotals2017-3.xlsx", 
                    range = "A2:B14", sheet = 3) %>%
  janitor::clean_names() %>%
  filter(!is.na(total)) %>%
  mutate(year = "2017")

trash_df_2016 = 
  readxl::read_xlsx("./data/HealthyHarborWaterWheelTotals2017-3.xlsx", 
                    range = "A2:B14", sheet = 4) %>%
  janitor::clean_names() %>%
  filter(!is.na(total)) %>%
  mutate(year = "2016")

# combine datasets and convert months
trash_bd = rbind(trash_df_2016, trash_df_2017) %>%
  mutate(month = month.name[month])

dim(trash_df)
dim(trash_bd)
trash_med = filter(trash_df ,year == "2016") 
  median(trash_med$sports_balls)
```
From the three different sheets, we can see the dimension of trash_df is 174,14, that of trash_bd is 15, 3; there are variables like year, month, total in trash_bd, or variables like sports_balls, weight_tons in trash_df. 
The total precipitaion in 2017 is 7.47. The median number of sports balls in a dumpster in 2016 is 26.

# Problem 3
```{r}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)

data("brfss_smart2010")

brfss = janitor::clean_names(brfss_smart2010) %>% 
  filter(topic == "Overall Health") %>%
  select(year, locationabbr, locationdesc, response, data_value) %>%
  spread(key = response, value = data_value) %>%
  janitor::clean_names() %>%
  mutate(exl_gd_prop = excellent + very_good )
  
  

```

## the questions
```{r (a)}
# unique locations
nrow(distinct(brfss,locationdesc))

nrow(distinct(brfss,locationabbr))
## there are 51 states here. Every states have been represented.
names(table(brfss$locationabbr))[which.max(table(brfss$locationabbr))]

## we can know that "NJ" is observed most.
```

```{r}
# get the median
brfss_2002 = filter(brfss, year == "2002") 
median(brfss_2002$excellent, na.rm = T)
```

## make a histogram
```{r}
ggplot(brfss_2002, aes(x = excellent)) + 
  geom_histogram() + ggtitle( "the histogram of Excellent") +
  ggsave("hist_Exl_2002.jpg",width = 8, height = 5)

```

## Make a scatterplot 
```{r}
# 
brfss %>% 
  filter( locationdesc == "NY - New York County"|locationdesc =="NY - Queens County") %>%
  ggplot(aes(x = year, y = excellent, color = locationdesc)) + 
  geom_point(alpha = .5, position = "dodge") +
  ggtitle( "the scatterplot of Excellent in different years") + 
  ggsave("points_Exl_yr.jpg",width = 8, height = 5)
```

