---
title: "p8105_hw2_cy2522"
author: "Chu YU"
date: "30th-Sep-2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# problem 1
## import and clean the data
```{r}
# import the csv data
library(tidyverse)
transit_df = 
  read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")%>%
  janitor::clean_names()

# clean the data
transit_tidy_df = 
  select( transit_df, line:entry, vending, entrance_type, ada)%>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

# dimension
dim(transit_tidy_df)
```
This dataset contains variable: line, station_name,station_latitude
,station_longitude, route_num ,route , entry, vending, entrance_type, ada. 
My data cleaning includes three steps: cleaning variables' names, deleting the unnecessary columns and removing the missing values.
The dimension of the dataset is 20548*10. They are not tidy.There are different routes in key variables and missing values.

## distinct 
```{r}
# How many distinct stations?
sn_dis = 
  distinct(transit_tidy_df[c("station_name","line" ,"ada")]) 
nrow(sn_dis)

# How many stations with ada:
filter(sn_dis, ada == "TRUE") %>% nrow()

# prop of vending

prop.table(table(transit_tidy_df$vending))
```

## Reformat
```{r}
# gather the route columns
transit_tidy_df_rt = 
  gather(transit_tidy_df, key = route_num, value = route, route1:route11)

# how many stations serving A ?
filter(transit_tidy_df_rt, route == "A") %>% 
  distinct(station_name,line) %>%
  nrow()

# serving A as well as ADA compliance
filter(transit_tidy_df_rt, route == "A"& ada == "TRUE") %>% 
  distinct(station_name,line) %>%
  nrow()
```


# Problem 2
```{r}
# import and clean the data
trash_df = 
  readxl::read_xlsx("./data/HealthyHarborWaterWheelTotals2017-3.xlsx", 
                    range = "A2:N213", sheet = 1) %>%
  janitor::clean_names() %>%
  filter(!is.na(dumpster)) %>%
  mutate(sports_balls = round(sports_balls, digits = 0)) 

# Read and clean precipitation data for 2016 and 2017
trash_df_2017 = 
  readxl::read_xlsx("./data/HealthyHarborWaterWheelTotals2017-3.xlsx", 
                    range = "A2:B15", sheet = 3) %>%
  janitor::clean_names() %>%
  filter(!is.na(total)) %>%
  mutate(year = "2017")

trash_df_2016 = 
  readxl::read_xlsx("./data/HealthyHarborWaterWheelTotals2017-3.xlsx", 
                    range = "A2:B15", sheet = 4) %>%
  janitor::clean_names() %>%
  filter(!is.na(total)) %>%
  mutate(year = "2016")

# combine datasets and convert months
trash_bd = rbind(trash_df_2016, trash_df_2017) %>%
  mutate(month = month.name[month])



```

4